---
title: "Lab 3 - Introduction to RStudio and R Markdown. Presentations"
author: "Omowunmi Obadero"
date: "September 20, 2025"
output:
  html_document:
    toc: true
<<<<<<< HEAD
editor_options: 
  markdown: 
=======
  pdf_document:
    toc: true
editor_options:
  markdown:
>>>>>>> 2ff06f15526d6988832d133980a39778bbb95e59
    wrap: 72
---

<!-- This is a comment in R Markdown (and regular markdown, such as in Jupyter)-->

<!-- The dashes below are not part of the YAML block above, -->

<!-- they insert a horizontal line after the ToC -->

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In Lab 3, you will learn how to use, edit and create a R Markdown
document (like this one) using RStudio. You should follow the
instructions in this document to complete the assignment. Knit this
document to view the nicely rendered HTML, which can make it easier to
read the questions.

If you need help as you use R Markdown in this lab and others in the
future, consult the following resources:

-   [Cheat sheet](https://rmarkdown.rstudio.com/lesson-15.html)
-   [Home page with guides](https://rmarkdown.rstudio.com/docs/)
-   [Reference book](https://bookdown.org/yihui/rmarkdown/)

The below is a code chunk, but instead of using the `r` engine we're
creating and alert block that will make the question show up with a blue
background in the HTML output. Unfortunately, this creates and error
when exporting to PDF, so it can only be used for HTML.

## Submission Instructions

rubric={mechanics:3}

You receive mark for submitting your lab correctly, please follow these
instructions:

-   [Follow the general lab
    instructions](https://ubc-mds.github.io/resources_pages/general_lab_instructions/).
-   [Click here to view a description of the rubrics used to grade the
    questions](https://github.com/UBC-MDS/public/tree/master/rubric).
-   Push your `.Rmd` AND **all the files** you will create as part of
    the lab to your GitHub repository.
    -   The reason for pushing all the files is that `.Rmd` does not
        contain the rendered output from running the cells. If someone
        is checking out your work there needs to be an HTML file to view
        the output, so it is good to get into this habit. - `.ipynb`
        renders nicely on GitHub, which is why we did not include the
        HTML file for previous labs.
-   Upload a `.Rmd` version of your assignment to Gradescope.
-   Include a clickable link to your GitHub repo for the lab just below
    this cell (it should look something like this
    <https://github.ubc.ca/MDS-2022-23/DSCI_521_labX_yourcwl>).

<https://github.ubc.ca/mds-2025-26/DSCI_521_lab3_group17>

## Editing R Markdown documents

This document is called an R Markdown document. It is a literate code
document, similar to Jupyter notebooks where you can write code and view
its outputs. To start, let's set our working directory by creating a new
R Project for lab 4.

### Text and rendering R Markdown documents

In a R Markdown document any line of text not in a code chunk (like this
line of text) will be formatted using Markdown. Similar to JupyterLab,
you can also use HTML and LaTeX here to do more advanced formatting. To
run a code chunk, you can press the green play button in the top right
corner of the chunk.

#### Question 1

rubric={correctness:1}

As you work on this lab you will be rendering the document(s). Please
make sure all source and redered documents are in the repository.

I've talked in class about making sure you ignore rendered output, but
for this assignment we'll treat it as an exception becuause we want to
confirm that the documents are properly formatted and rendering.

Nothing to post. We will be looking at your final repository contents.

#### Question 2

rubric={mechanics:1}

Create a new code chunk below using the r language engine that runs some
R code (it does not need to be complicated, but it should have an
output). Ensure that you can render/knit the document after you add that
chunk.

```{r value of squared e}
x <- exp(1)^2
print(round(x, 2))
```

#### Question 3

rubric={mechanics:1}

Create a new code chunk, and add a meaningful name to the code chunk.
Try using the pop-up-like menu to navigate between the named code chunks
Don't forget to knit/render the document after you make this change to
ensure everything is still working.

```{r calculate average score}
math_score <- 89
stat_score <- 92
cs_score <- 70

mean_score <- mean(c(math_score, stat_score, cs_score))
print(round(mean_score, 2))
```

#### Question 4

rubric={mechanics:1,reasoning:1}

Create a new code chunk that uses a code chunk option. Write out in your
own words what the code chunk option is doing.

Added the option echo = FALSE which hides the code chunk and only
displays the result (default is TRUE)

<<<<<<< HEAD
#### Multiple code chunk options

=======
```{r echo = FALSE}
my_var <- "Hi! this is a surprise string"
print(my_var)
```

#### Multiple code chunk options

>>>>>>> 2ff06f15526d6988832d133980a39778bbb95e59
To have multiple code chunk options you separate them by a comma. For
example, if in addition to suppressing warnings, we want to run the code
but not output the results, then we can add the `include = FALSE`
argument to the code chunk after the `warning = FALSE` option.

#### Question 5

rubric={mechanics:1,reasoning:1}

Create a new code chunk that uses at least two code chunk options. At
least one must be different to the ones mentioned above. Write in your
own words what each code chunk option is doing.

ADDED 2 options

-   **warning = FALSE** makes the output not display any warnings
    associated with running the chunk for

    this case the text "G2;H2;Warningh: [38;5;254mRemoved 2 rows
    containing missing values or values outside the scale range
    (`geom_point()`).[39mg" is removed if we add warning = FALSE

-   **fig.width = 5** sets the width of the output figure to 5 (inches)
    (default is 7)

```{r warning = FALSE, fig.width = 5}
library(ggplot2)
library(palmerpenguins)
my_plot <- ggplot(data = penguins, mapping = aes(x = bill_length_mm, 
                                      y = body_mass_g,
                                     colour = species)) +
    geom_point(size = 2) +
    xlab("Bill length (mm)") +
    ylab("Body mass (g)") +
    theme_bw() +
    theme(text = element_text(size = 18))
my_plot
```

### YAML Header and document output options

R Markdown files contains three types of content:

1.  Plain text mixed with simple Markdown formatting.
2.  Code chunks surrounded by <code>\`\`\`</code>.
3.  An (optional) YAML header surrounded by `---`.

You have been introduced the first two types of content, but not the
third (although you probably saw it at the top of this document). The
(optional) YAML header, which is located at the very top of R Markdown
files sets some general global parameters, including:

-   title
-   author
-   output
-   etc

**Example YAML Header**

```         
---
title: "Reproducible Data Science Report"
author: "Florencia D'Andrea"
date: "September 4, 2022"
output: html_document
---
```

Most important from a workflow perspective is **output**. Possible
output options include:

-   `output: html_document`
-   `output: md_document`
-   `output: pdf_document`
-   `output: word_document`
-   `output: xaringan::moon_reader` (xaringan presentation - html)

#### Question 6

rubric={mechanics:1}

Navigate to the YAML header at the very top of this document and edit it
so that you include an `author` (yourself) and a `date` (lab due date).
Include what you added below here as well as a fenced Markdown code
block.

```{}
author: "Omowunmi Obadero"
date: "September 20, 2025"
```

### Creating R Markdown documents

You can use the "File" menu inside RStudio to create new R Markdown
documents by selecting: **File \> New File \> R Markdown** This will
bring you to another menu where you can choose the type of output (don't
be afraid to pick something, you can always change the `output` type
once you have the `.Rmd` file).

To create a written report, we generally recommend using the default
`output: html_document` as it is easier to read than PDF (note - LaTeX
does not render nicely in such documents sadly, so if you are using a
lot of LaTeX then you may want to choose `output: pdf_document`). If you
want to create an `.md` file to publish on GitHub, it is recommend to
instead use `output: github_document`. To get this from the menu above
you need to navigate to the "From Template" option on the left panel and
then select "GitHub Document (Markdown)".

#### Question 7

rubric={mechanics:2}

1 - Create a new RMarkdown report (a different file than this one) in
the same directory as this RMarkdown file. Use `html_document` as the
`output`. After you have rendered it, paste the link to the HTML output
as a link to your GitHub repository (remember to push all your files!)

2 - Then, navigate to the YAML header at the very top of that `.Rmd`
document and edit it so that the `output` is `pdf_document`. Then
knit/render the document. Note the different output. Add and commit that
rendered both the `.html` and `.pdf` files to the GitHub repository for
this lab and paste the two links below this question.

**7.1** html link:
<https://github.ubc.ca/mds-2025-26/DSCI_521_lab3_group17/blob/master/q7.html>

**7.2** pdf link:
<https://github.ubc.ca/mds-2025-26/DSCI_521_lab3_group17/blob/master/q7.pdf>

#### Question 8

rubric={mechanics:6}

1.  Go back to the `.Rmd` file you created in question 7, and include at
    least two Markdown text sections (each should have a header) and at
    least two separate code chunks in it (these can be really simple).
    Save the new R Markdown document and give it a new meaningful name.

2.  Render/knit the new R Markdown document to get an `.html` file. Put
    the `.Rmd` document and the rendered `.html` file under version
    control using Git, and push/upload the file to your GitHub
    repository for this homework. Paste a link to these files as your
    answer below.

**8**

Rmd link:
<https://github.ubc.ca/mds-2025-26/DSCI_521_lab3_group17/blob/master/q8.Rmd>

html link:
<https://github.ubc.ca/mds-2025-26/DSCI_521_lab3_group17/blob/master/q8.html>

#### Question 9 (Optional)

rubric={mechanics:1,reasoning:1}

1.  Take the R Markdown report created in Question 8 and change the
    output to `github_document` and render it. Put the rendered `.md`
    file under version control using Git, and push/upload the file to
    your GitHub repository for this homework. Try to look at the file on
    GitHub.ubc.ca in your homework repo? What do you see? How is it
    rendered?

The markdown document is automatically rendered to be a human readable
webpage unlike the html file. The markdown document is rendered by its
own markdown engine through GitHub Flavored Markdown syntax.

#### Question 10

rubric={mechanics:6}

1.  Create a presentation using RStudio. Do this in a different file
    than this one but in the same directory as this RMarkdown file. You
    can use xaringan or Quarto to create the slides. [On the
    book](https://pages.github.ubc.ca/MDS-2023-24/DSCI_521_platforms-dsci_students/lectures/6-rmarkdown-quarto-slides-ghpages.html)
    you will find links that will guide you on how to create each type
    of slide. We will accept both Xaringan, Quarto, or RISE slides as
    correct. Give this file a meaningful name.
2.  Create at least 4 slides. At least two slides must include a code
    chunk or cell (these can be really simple). Save the new document.
3.  Render/knit/export the new document to get a `html` presentation
    file.
4.  Put the new document and the rendered `.html` file under version
    control using Git, and push/upload the file to your GitHub
    repository for this lab
5.  Activate GitHub pages and paste the link below. Remember where
    github looks for website materials, and make sure you put the files
    in the approriate directory. You can use any of the folder/branches
    to publish the slides, as long as the URL works.

YOUR ANSWER GOES HERE

#### (Challenging) Question 11

rubric={reasoning}

In a paragraph or two, compare and contrast the use of reproducible
tools (e.g., R Markdown and Jupyter) and non-reproducible tools (Word,
Powerpoint, Keynote, etc) for presentations and reports. Include
advantages and disadvantages for each.

**Reproducible tools**:

- Advantages: 

(1) The data, analysis code, and output are all 
integrated together, which means they are literate programming documents. So the whole presentation/report can be 
fully reproduced as long as we have the source files, and the files are very readable and dynamic since they contain
both the code and the explaining text. (2) Outputs can be rendered into multiple formats easily, such as `html`, `pdf`, etc.
(3) Version control can be easily embedded, facilitating collaboration among authors.

- Disadvantages: 

(1) It is less user-friendly. It actually requires knowledge and training of version control and coding.
(2) There could be potential merge conflicts, which makes collaboration a bit more difficult than live editing.

**Non-reproducible tools**:

- Advantages: 

(1) It is more user-friendly in the sense that it is easier to create and modify the presentation/report.
In other words, users do not have to have deep understanding of coding and version control. So these tools are a great
start for beginners. (2) There are more templates published, so using these tools can produce a more visual-appealing 
presentation/report. (3) Some of these tools allow simultaneous editing of multiple authors, and the results are updated
instantaneously, without having to go through the complicated version control system.

- Disadvantages: 

(1) Source code cannot be easily integrated. Every time the data is updated, we have to re-create
the presentation/report again to update the values manually. (2) It cannot track different versions of the presentation/report
in the history as the version control does. So we cannot revert back to a previous version if we want to. (3) The options of 
output rendering are limited.

## Free Point

Set `free_point` to `TRUE` with `free_point <- TRUE`

```{r tags=c()}
free_point <- TRUE
```

```{r}
. = ottr::check("tests/free-point-autograde.R")
```

## Submission

Make sure you have run all cells in your notebook in order before
running the cell below, so that all images/graphs appear in the output.
The cell below will generate a zip file for you to submit. **Please save
before exporting!**

```{r}
# Save your notebook first, then run this cell to export your submission.
# ottr::export("lab3.Rmd", pdf = TRUE)
```
